---
output: pdf_document
---
#Reproducible Research - Course Project 1


##Step 1 : Loading the data

```{r pre-processing}
setwd("c:/users/vijay/documents/rep_research/data")
activity = read.csv("activity.csv")
```


##Step 2 : What is the mean total number of steps taken per day?

###Histogram of the total number of steps taken each day

```{r code to create the histogram}
#use tapply to sum the number of steps by date
steps_by_date = as.numeric(tapply(activity$steps, activity$date, sum, na.rm = T))
#plot a histogram
hist(steps_by_date, main = "Total Number of Steps Taken Each Day")
```

###Mean and Median of the Total Number of Steps Taken Each Day
```{r mean and median}
#mean of the number of steps taken each day
mean(steps_by_date)
#median of the number of steps taken each day
median(steps_by_date)
```


##Step 3 : What is the average daily activity pattern?

###Time-Series plot of daily activity pattern 

```{r average daily activity pattern}
adap = as.numeric(tapply(activity$steps, activity$interval, mean, na.rm = T))
plot(adap, type="l", xlab = "Interval", ylab = "Number of Steps", main = "Average Daily Activity Pattern")
```

###5-minute interval with the maximum number of steps

```{r interval with max steps}
which.max(adap)
```

####The 104th 5-minute interval (approximately 9 am) has the most number of steps.


##Step 4 : Imputing Missing Values

###Total Number of rows with missing values (NAs)

```{r number of missing value rows}
table(is.na(activity))
```

####There are 2,304 rows with missing values in the "activity"" data frame (all from the "steps" field)

###Create dataset with missing values filled by imputed values

```{r create dataset with imputed values}
#use "mice" package
library(mice)
#set the seed so as to make it reproducible
set.seed(144)
#do the imputation to create the new dataset("activity2") with the missing data filled in
activity2 = complete(mice(activity))
```

####Histogram of the total number of steps taken each day (including imputed data)

```{r code to create the imputed values histogram}
#use tapply to sum the number of steps by date
steps_imputed = as.numeric(tapply(activity2$steps, activity$date, sum, na.rm = T))
#plot a histogram
hist(steps_imputed, main = "Total Number of Steps Taken Each Day")
```

###Mean and Median of the Total Number of Steps Taken Each Day
```{r mean and median for imputed dataset}
#mean of the number of steps taken each day
mean(steps_imputed)
#median of the number of steps taken each day
median(steps_imputed)
```

####The histogram is almost the same. Mean and median values, expectedly increase, as additional values have been added.





